The results from the NSGA-II optimization process reveal the complexity of the Electric Vehicle Charging Station (EVCS) deployment problem. By considering five key objectives—maximizing coverage, maximizing charger speed, minimizing the number of stations, minimizing the number of chargers, and minimizing the average distance between electric vehicles (EVs) and their nearest charging station—the optimization process provides a set of Pareto-optimal solutions that represent the trade-offs between these conflicting objectives. The use of NSGA-II enabled the exploration of the solution space, resulting in a diverse range of solutions that balance infrastructure cost, service quality, and accessibility.

The Pareto front obtained from the optimization illustrates that there is no single best solution. Instead, decision-makers can select the optimal solution based on their specific priorities, whether it be improving coverage, reducing infrastructure costs, or enhancing service accessibility. This flexibility highlights the importance of considering all relevant factors when planning the deployment of EV charging infrastructure.

Through the optimization process, we achieved a significant reduction in infrastructure costs, with the number of stations decreasing from 92 to 66 and the number of chargers reducing from 91 to 81, while maintaining a high level of service. Optimizing station locations ensured optimal coverage, and reducing the number of stations and chargers per station helped reduce associated costs. The trade-offs identified in the results underscore the need for careful planning in the deployment of EVCS to ensure an effective balance between coverage, cost, and accessibility.

In conclusion, the application of NSGA-II demonstrates an effective approach for optimizing electric vehicle charging station infrastructure planning by addressing the research question of how multiobjective evolutionary algorithms can balance conflicting objectives in EVCS deployment. The results indicate that this method can support the development of cost-effective and efficient charging networks, providing electric vehicle users with reliable access to quality charging services while managing infrastructure costs. This approach offers a solid foundation for further studies and practical implementations aimed at advancing sustainable transportation systems.

