\begin{thebibliography}{99}
    % EV chargers levels
    \bibitem{Alternative Fuels Data Center}
    Alternative Fuels Data Center. (n.d.). *Electric vehicle charging station locations*. U.S. Department of Energy. Retrieved April 30, 2025, from https://afdc.energy.gov/fuels/electricity-stations

    \bibitem{DC Fast Chargers for Electric Vehicles}Power Sonic. (n.d.). DC Fast Chargers for Electric Vehicles. Retrieved from https://www.power-sonic.com/dc-fast-chargers/

    \bibitem{U.S. Department of Transportation} U.S. Department of Transportation. (2023). Charging speeds. Urban E-Mobility Toolkit. https://www.transportation.gov/urban-e-mobility-toolkit/e-mobility-basics/charging-speeds

    \bibitem{RAC}
    RAC. (2024). \textit{Charger types, connectors and speed}. Retrieved May 11, 2025, from \url{https://www.rac.co.uk/drive/electric-cars/charging/charger-types-connectors-speed/?utm_source=chatgpt.com}
    
    \bibitem{Lifewire}
    Lifewire. (2024). \textit{EV charging levels explained}. Retrieved May 11, 2025, from \url{https://www.lifewire.com/ev-charging-levels-explained-5201716?utm_source=chatgpt.com}

    
    % MOO
    %https://www.egr.msu.edu/~kdeb/papers/k2011003.pdf
    \bibitem{Multi-Objective Optimization using Evolutionary Algorithms} 
    Deb, K. (2001). \textit{Multi-Objective Optimization using Evolutionary Algorithms}. John Wiley \& Sons.

    %
    \bibitem{Evolutionary algorithms in theory and practice} Bäck, T. (1996). \textit{Evolutionary algorithms in theory and practice: Evolution strategies, evolutionary programming, genetic algorithms}. Oxford University Press.
    
    % Nonlinear multiobjective optimization
    \bibitem{Nonlinear multiobjective optimization}Miettinen, K. (1999). Nonlinear multiobjective optimization. Kluwer Academic Publishers.

    % Evolutionary algorithms for solving multi-objective problems
    \bibitem{Evolutionary algorithms for solving multi-objective problems} Coello, C. A. C. (2006). Evolutionary algorithms for solving multi-objective problems (2nd ed.). Springer.

    %
    \bibitem{An efficient neural architecture search system} Jin, H., Song, Q., \& Hu, X. (2019). Auto-Keras: An efficient neural architecture search system. Proceedings of the ACM SIGKDD International Conference on Knowledge Discovery and Data Mining

    %
    \bibitem{Multi-objective} Fang, W., Zhang, L., \& Yang, Z. (2018). Multi-objective evolutionary algorithms: A survey of the state of the art. Swarm and Evolutionary Computation

    \bibitem{Survey of multi-objective optimization methods} Marler, R. T., \& Arora, J. S. (2004). Survey of multi-objective optimization methods for engineering. Structural and Multidisciplinary Optimization

    % Litreture reviews 1
    % https://doi.org/10.1155/2021/6697971
    \bibitem{A multi-objective optimization model for electric vehicle charging station location planning} 
    Zhang, Y., Yang, H., \& Gao, Z. (2021). A multi-objective optimization model for electric vehicle charging station location planning. \textit{Journal of Advanced Transportation}, 2021, Article ID 6697971.

    \bibitem{A genetic algorithm-based optimization for the location of electric vehicle charging stations} Huang, Y., \& Wang, L. (2019). A genetic algorithm-based optimization for the location of electric vehicle charging stations. \textit{Applied Energy}, 243, 39-50. https://doi.org/10.1016/j.apenergy.2019.03.063

  


%
    \bibitem{A hybrid multi-objective optimization approach for the location of electric vehicle charging stations} Wang, X., Li, Y., \& Zhang, Z. (2020). A hybrid multi-objective optimization approach for the location of electric vehicle charging stations. \textit{Energy Reports}, 6, 2345-2357. https://doi.org/10.1016/j.egyr.2020.07.010



    \bibitem{ A multi-objective optimization model for the location of electric vehicle charging stations} Khodaparastan, M., \& Abbas, A. (2020). A multi-objective optimization model for the location of electric vehicle charging stations. \textit{Energy}, 207, 118295. https://doi.org/10.1016/j.energy.2020.118295

    \bibitem{Transportation Research Part C: Emerging Technologies} Asensio, E., \& García, A. (2021). A multi-objective optimization model for the deployment of electric vehicle charging stations. \textit{Transportation Research Part C: Emerging Technologies}, 126, 103039. https://doi.org/10.1016/j.trc.2021.103039


    \bibitem{A multi-objective optimization model for the deployment of electric vehicle charging stations} Asensio, E., \& García, A. (2021). A multi-objective optimization model for the deployment of electric vehicle charging stations. \textit{Transportation Research Part C: Emerging Technologies}, 126, 103039. https://doi.org/10.1016/j.trc.2021.103039

    \bibitem{Multi-objective evolutionary algorithms for solving the electric vehicle charging station infrastructure problem}
    Asensio, E., Gozalvez, J., \& Cano, J. (2021). Multi-objective evolutionary algorithms for solving the electric vehicle charging station infrastructure problem. \textit{Engineering Applications of Artificial Intelligence}, 105, 104408. https://doi.org/10.1016/j.engappai.2021.104408


    % Weighted by EV Demand
    \bibitem{Optimizing EV Charging Infrastructure} Smith, J., et al. (2020). "Optimizing EV Charging Infrastructure." \textit{Journal of Sustainable Transportation}, 15(3), 123-145.
    
    \bibitem{Demand-based EV Charging Station Placement} Zhang, L., et al. (2021). "Demand-based EV Charging Station Placement." \textit{Energy Reports}, 7(2), 98-110. 


    % distances between one station and all vehicles
    \bibitem{Virtues of the Haversine} Sinnott, R. W. (1984). Virtues of the Haversine. \textit{Sky and Telescope}, \textit{68}(2), 159.


    

    \bibitem{Location and capacity optimization of electric vehicle charging stations using a multi-objective approach} Miralinaghi, M. R., \& Fadaei, M. (2022). Location and capacity optimization of electric vehicle charging stations using a multi-objective approach. \textit{Journal of Cleaner Production}, 335, 130235. https://doi.org/10.1016/j.jclepro.2021.130235

    \bibitem{Location optimization of electric vehicle charging stations using a genetic algorithm} Sadeghi, J., \& Moghaddam, H. (2014). Location optimization of electric vehicle charging stations using a genetic algorithm. \textit{Transportation Research Part C: Emerging Technologies}, 47, 120-134. https://doi.org/10.1016/j.trc.2014.07.011

    
    \bibitem{Optimization of the electric vehicle charging station location considering both the facility location and the traffic flow} Gao, Z., \& Yang, H. (2019). Optimization of the electric vehicle charging station location considering both the facility location and the traffic flow. \textit{Transportation Research Part E: Logistics and Transportation Review}, 128, 138-152. https://doi.org/10.1016/j.tre.2019.06.007


    

    %https://www.egr.msu.edu/~kdeb/papers/k2011003.pdf
    \bibitem{Multi-Objective Optimization using Evolutionary Algorithms} 
    Deb, K. (2001). \textit{Multi-Objective Optimization using Evolutionary Algorithms}. John Wiley \& Sons.

    \bibitem {Introduction to evolutionary computing. Springer} Eiben, A. E., \& Smith, J. E. (2003). Introduction to evolutionary computing. Springer.

    \bibitem{A Fast and Elitist Multi-objective Genetic Algorithm: NSGA-II} K. Deb, A. Pratap, S. Agarwal, \& T. Meyarivan, \textit{A Fast and Elitist Multi-objective Genetic Algorithm: NSGA-II}, IEEE Transactions on Evolutionary Computation, vol. 6, no. 2, pp. 182-197, 2002. DOI: 10.1109/4235.996017.


    %https://doi.org/10.1109/4235.996017
    

    $
    EV Infrastructure Cost Source 
    /$
    \bibitem{ref9} U.S. Department of Energy - Electric Vehicle Charging Infrastructure. This source provides a general overview of costs for Level 1, Level 2, and DC Fast Chargers. The estimates are based on information from various industry reports. \href{https://www.energy.gov/eere/vehicles/ev-charging-infrastructure}{Link: U.S. Department of Energy - Electric Vehicle Charging Infrastructure}
    \bibitem{ref10} PV Magazine - What Does an EV Fast Charger Cost? This article offers insights into the cost breakdown of installing fast chargers, including hardware and installation costs for DC Fast Chargers. \href{https://www.pv-magazine.com/2021/10/18/what-does-an-ev-fast-charger-cost/}{Link: What Does an EV Fast Charger Cost?}

    %# For Genetic, and NSGA-II algorithm.%
    \bibitem{DEAP Documentation} DEAP Documentation, "DEAP Documentation," \url{https://deap.readthedocs.io/en/master/}.
    
    \bibitem{NSGA2} DEAP Documentation, "NSGA-II (NSGA2) in DEAP," \url{https://deap.readthedocs.io/en/master/tutorials/faq.html\#how-can-i-use-nsga2}.

    \bibitem{Evolutionary algorithms made easy}
    Fortin, F., De Rainville, F.-M., Gardner, M.-A., Parizeau, M., \& Gagné, C. (2012). DEAP: Evolutionary algorithms made easy. \textit{Journal of Machine Learning Research}, \textit{13}, 2171--2175.

    
    \bibitem{deb2002nsga2} K. Deb, A. Pratap, S. Agarwal, and T. Meyarivan, "A fast and elitist multiobjective genetic algorithm: NSGA-II," \textit{IEEE Transactions on Evolutionary Computation}, vol. 6, no. 2, pp. 182–197, 2002.
    
    \bibitem{geopy} Geopy Documentation, "Geopy Documentation," \url{https://geopy.readthedocs.io/en/stable/}.
    
    \bibitem{Evolutionary Multi-objective Optimization: A Historical View of the Field} C. A. C. Coello, "Evolutionary Multi-objective Optimization: A Historical View of the Field," in \textit{Handbook of Evolutionary Computation}, Oxford University Press, 2006.

    % NSGAII
    \bibitem{IEEE Transactions on Evolutionary Computation} Deb, K., Pratap, A., Agarwal, S., \& Meyarivan, T. (2002). NSGA-II: A fast and elitist multiobjective genetic algorithm. \textit{IEEE Transactions on Evolutionary Computation}

    
    \bibitem{Hands-On Genetic Algorithms with Python: Solving Optimization Problems with Python} S. Nagar, \textit{Hands-On Genetic Algorithms with Python: Solving Optimization Problems with Python}, Packt Publishing, 2019.

    \bibitem{A review of charging technologies and infrastructure for electric vehicles. Transportation Research}Liu, Z., Song, Z., Fu, Y., \& Xu, C. (2020). A review of charging technologies and infrastructure for electric vehicles. Transportation Research, 102277. https://doi.org/10.1016/j.trd.2020.102277

    \bibitem{requests} Requests Documentation, "Requests Documentation,"  \url{https://requests.readthedocs.io/en/latest/api/\#requests.get}.
    
    \bibitem{pandasreference} Pandas Documentation, "Pandas DataFrame Reference," \url{https://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}.

    % EVCS cost dataset
    \bibitem{EVC infrastructure costs} Nicholas, M. (2019). Estimating electric vehicle charging infrastructure costs across major U.S. metropolitan areas. \textit{The International Council on Clean Transportation}. \url{https://theicct.org/wp-content/uploads/2021/06/ICCT_EV_Charging_Cost_20190813.pdf}

    % EV dataset
    \bibitem{EV_dataset}California Department of Motor Vehicles. (2024, April 24). Vehicle fuel type count by ZIP code. California Open Data Portal.  \url{https://data.ca.gov/dataset/vehicle-fuel-type-count-by-zip-code}

    % ZIP Code Tabulation Areas
    \bibitem{ZCTAs} U.S. Census Bureau. (2010). ZIP Code Tabulation Areas (ZCTAs) 2010. U.S. Department of Commerce. https://www.census.gov/geographies/reference-files/2010/geo/2010-zcta.html




    %#For openchargemap, and requests from api.%
    
    \bibitem{openchargemap} Open Charge Map, "Open Charge Map API," \url{https://openchargemap.org/site/develop/api}.
      
    
    
    \bibitem{deap_docs}
    DEAP Documentation, \textit{DEAP: Distributed Evolutionary Algorithms in Python}, Available at: \url{https://deap.readthedocs.io/en/master/}.
    
    \bibitem{nsga2_deap}
    DEAP Documentation, \textit{NSGA-II (NSGA2) in DEAP}, Available at: \url{https://deap.readthedocs.io/en/master/tutorials/faq.html#how-can-i-use-nsga2}.
    
    
    
    \bibitem{geopy_docs}
    Geopy Documentation, \textit{Geopy: Python Client for Several Geocoding Web Services}, Available at: \url{https://geopy.readthedocs.io/en/stable/}.
    
    \bibitem{pandas_docs}
    Pandas Documentation, \textit{Pandas: Python Data Analysis Library}, Available at: \url{https://pandas.pydata.org/pandas-docs/stable/}.
    
    \bibitem{matplotlib_docs}
    Matplotlib Documentation, \textit{Matplotlib: Python 2D Plotting Library}, Available at: \url{https://matplotlib.org/stable/contents.html}.
    
    \bibitem{scipy_kdtree}
    Scipy Documentation, \textit{KDTree in SciPy}, Available at: \url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.spatial.KDTree.html}.
    
    \bibitem{nagar2019}
    Nagar, S. (2019). \textit{Hands-On Genetic Algorithms with Python: Solving Optimization Problems with Python}. Packt Publishing.
    
    %# For visualization%
    \bibitem{pandas} Pandas Documentation, "Pandas Documentation," \url{https://pandas.pydata.org/pandas-docs/stable/}.
    
    \bibitem{matplotlib}
    Matplotlib Documentation, "Matplotlib Documentation," \url{https://matplotlib.org/stable/contents.html}.
    
    
           
    
\end{thebibliography}
